\( \tool \): JavaScript IR-based Semantics Extraction Toolchain

### 2020-04-03 ase20/overview.tex
Revised Sec. 2

In this section, we introduce the overall structure of \( \tool \)
depicted in Figure~\ref{fig:overview} and the outline of the paper.
Compared to the existing approaches shown in
Figure~\ref{fig:existing}, \( \tool \) automatically synthesizes
{\sf JavaScript Parser} and {\sf AST-\( \ires \) Translator} directly from
ECMAScript.

The motivation of automatically extracting syntax and semantics of JavaScript
from the ECMAScript is twofold: 1) ECMAScript is written in a well-organized
style, and 2) it is available as HTML files using various HTML tags.  As the
first step, we developed {\sf Spec Extractor} to simply collect and save both
syntax and semantics into JSON format.  We implemented {\sf Spec Extractor} in
JavaScript to easily manipulate ECMAScript written in HTML.

\begin{figure}
  \centering
  \begin{subfigure}[t]{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/arrayliteral-syntax.pdf}
    \caption{The \textit{ArrayLiteral} production in ES10}
    \label{fig:array-literal-es}
  \end{subfigure}
  \begin{subfigure}[t]{0.48\textwidth}
    \begin{lstlisting}[style=myScalastyle]
val ArrayLiteral: List[Boolean] => LAParser[T] = memo {
  case List(Yield, Await) =>
  "[" ~ opt(Elision) ~ "]"             ^^ ArrayLiteral0 |
  "[" ~ ElementList(Yield,Await) ~ "]" ^^ ArrayLiteral1 |
  "[" ~ ElementList(Yield,Await) ~ ","
      ~ opt(Elision)~ "]"              ^^ ArrayLiteral2
}
    \end{lstlisting}
    \vspace*{-1em}
    \caption{The generated parser for the \textit{ArrayLiteral} production}
    \label{fig:array-literal-parser}
  \end{subfigure}
  \caption{The \textit{ArrayLiteral} production in the ES10 and its generated parser}
  \label{fig:array-literal}
  \vspace*{-1em}
\end{figure}

@@

To generate {\sf JavaScript Parser} from a given \( \bnfes \) grammar, we
construct \textsf{Parser Generator} in Scala.  It synthesizes a
JavaScript parser according to the given \( \bnfes \), and the
generated parser is defined with Scala parser combinators~\cite{scala-parser-combinators}.
Moreover, in order to parse \( \bnfes \) grammars correctly and
efficiently, we propose \textit{lookahead parsers}, which keep track
of lookaheads, sets of possible next tokens.  With lookahead parsers,
generated parsers become more readable and have one-to-one
mapping to their corresponding grammar productions.
For example, Figure~\ref{fig:array-literal}(b) shows the generated
parser for the \textit{ArrayLiteral} production in Figure~\ref{fig:array-literal}(a).
Each parser has the \( \code{List[Boolean] => LAParser[T]} \) type
because each production in \( \bnfes \) is parametric with boolean values.
The \( \code{memo} \) is a memoization function for pairs of boolean
parameters and resulting parsers for performance optimization.
The lazy value \( \code{ArrayLiteral} \) corresponds to the
\textit{ArrayLiteral} production.  In the parser, each string literal
such as \( \code{"["} \) or \( \code{"]"} \) denotes a parser for a
terminal symbol.  The \( \code{opt} \) helper function creates
optional parsers.  The parametric non-terminal \textit{ElementList}
with arguments \textsf{Yield} and \textsf{Await} is represented as a
function call \( \code{ElementList(Yield, Await)} \).
The \( \code{\textasciitilde} \) operator combines two parsers
and the \( \code{\^{}\^{}} \) operator describes how to construct ASTs.
When the left-hand side of \( \code{\^{}\^{}} \) is matched, its
right-hand side shows a corresponding AST constructor, where each
constructor has a number denoting the order among alternatives.
For example, the \( \code{ArrayLiteral0} \) constructor corresponds to
the first alternative of the \textit{ArrayLiteral} production.

@@

\textbf{Semantics.}
ECMAScript describes the language semantics as abstract algorithms in English.
While they are written in a natural language, the writing style is
well-organized with ordered steps and tagged tokens.  Our {\sf Spec Extractor}
reads each abstract algorithm with HTML tags and converts them into JSON files.
For example, Figure~\ref{fig:array-literal-eval}(a) presents the
\textbf{Evaluation} abstract algorithm of the third alternative of the
\textit{ArrayLiteral} production in ES10, and it has seven steps.  In the HTML
files describing the abstract algorithms, each non-terminal symbol
(i.e. \textit{ElementList}), local variable (i.e. \textit{array}), code (i.e. \(
\code{"length"} \)), or value (i.e. \textbf{false}) has the \( \code{<nt>} \),
\( \code{<var>} \), \( \code{<code>} \), or \( \code{<emu-val>} \) tag,
respectively.

To translate such abstract algorithms into representations that are
suitable for manipulation, we define \( \ires \), a specialized
intermediate representation for ECMAScript abstract algorithms.  Then,
we develop \textsf{Algorithm Compiler} in Scala using Scala parser
combinators again to convert given abstract algorithms to \( \ires \)
functions.  It also takes \textsf{Compile Rules} as another input,
which has two parts: parsing rules and conversion rules.  They are
\textit{manually} specified for ECMAScript; we explain them in
detail in Section~\ref{sec:compiler}.  Thus, each abstract algorithm is
converted into a function written in \( \ires \) via \textsf{Algorithm Compiler}.
For example, Figure~\ref{fig:array-literal-eval}(b) presents the generated
\( \ires \) function for the \textbf{Evaluation} abstract algorithm described in
Figure~\ref{fig:array-literal-eval}(a).
The \( \code{ArrayLiteral[2].Evaluation} \) function takes two parameters for
two non-terminal symbols: \( \code{ElementList} \) and \( \code{Elision} \).
The parameter \( \code{Elision} \) has a special value \( \code{absent} \) when
the non-terminal symbols \textit{Elision}$_{\mbox{\footnotesize opt}}$ is not
present.  Thus, we convert the condition in the step 4,
``if \textit{Elision} is not present,'' into the equality check with \(
\code{absent} \): \( \code{if (= Elision absent)} \).  Codes are represented as
string values and values are represented as corresponding \( \ires \) values.
For instance, the code \texttt{"length"} and the value \textbf{false} are
converted into the string value \( \code{"length"} \) and the boolean value
\( \code{false} \), respectively.

Finally, \( \tool \) constructs {\sf AST-\( \ires \) Translator} with the given
\( \ires \) functions and \textit{manually} specified {\sf Global Setting},
which has minor but necessary information to evaluate JavaScript programs
described in ECMAScript.  They consists of the structure of the standard
built-in objects and ECMAScript data types. Putting them all together, we can
translate a given JavaScript program into \( \ires \) via generated {\sf
JavaScript Parser} and {\sf AST-\( \ires \) Translator} by \( \tool \).  Even
though \( \tool \) is not fully automatic because of {\sf Compile Rules} and
{\sf Global Setting}, it could dramatically reduce the efforts to building
parsers and translators from the scratch.

\textbf{Outline.}
In the remainder of this paper, we describe the detailed techniques of
\( \tool \).  We first formally define \( \bnfes \) and how to
automatically generate parsers from given \( \bnfes \) using lookahead
parsers (Section~\ref{sec:parser}).  Then, we explain how to compile
abstract algorithms with tagged tokens into \( \ires \) functions using
compile rules (Section~\ref{sec:compiler}).  After describing implementation
details (Section~\ref{sec:impl}), we evaluate our tool \( \tool \)
(Section~\ref{sec:eval}), discuss related work (Section~\ref{sec:related}), and
conclude (Section~\ref{sec:conclude}).

>>>>

In this section, we introduce the overall structure of \( \tool \)
depicted in Figure~\ref{fig:overview}.
Compared to the existing approaches shown in
Figure~\ref{fig:existing}, \( \tool \) automatically synthesizes
{\sf JavaScript Parser} and {\sf AST-\( \ires \) Translator} directly from
ECMAScript.

The motivation of automatically extracting syntax and semantics of JavaScript
from ECMAScript is twofold: 1) ECMAScript is written in a well-organized
style, and 2) it is available as HTML files using various HTML tags.
To easily manipulate ECMAScript written in HTML, we developed
{\sf Spec Extractor} in JavaScript to extract the syntax and semantics in JSON format.

\begin{figure}
  \centering
  \begin{subfigure}[t]{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/arrayliteral-syntax.pdf}
    \caption{\textit{ArrayLiteral} production in ES10}
    \label{fig:array-literal-es}
  \end{subfigure}
  \begin{subfigure}[t]{0.48\textwidth}
    \begin{lstlisting}[style=myScalastyle]
val ArrayLiteral: List[Boolean] => LAParser[T] = memo {
  case List(Yield, Await) =>
  "[" ~ opt(Elision) ~ "]"             ^^ ArrayLiteral0 |
  "[" ~ ElementList(Yield,Await) ~ "]" ^^ ArrayLiteral1 |
  "[" ~ ElementList(Yield,Await) ~ ","
      ~ opt(Elision)~ "]"              ^^ ArrayLiteral2
}
    \end{lstlisting}
    \vspace*{-1em}
    \caption{Generated parser for the \textit{ArrayLiteral} production}
    \label{fig:array-literal-parser}
  \end{subfigure}
  \vspace*{-1em}
  \caption{\textit{ArrayLiteral} production in ES10 and its parser}
  \label{fig:array-literal}
\end{figure}

@@

To generate {\sf JavaScript Parser} from a given \( \bnfes \) grammar, we
construct \textsf{Parser Generator} in Scala.  It synthesizes a
JavaScript parser according to the given \( \bnfes \), and the
generated parser is defined with Scala parser combinators~\cite{scala-parser-combinators}.
Moreover, in order to parse \( \bnfes \) grammars correctly and
efficiently, we propose \textit{lookahead parsers}, which keep track
of lookaheads, sets of possible next tokens.  With lookahead parsing,
generated parsers now have one-to-one mapping to their corresponding
grammar productions, improving readability.
For example, Figure~\ref{fig:array-literal}(b) shows the generated
parser for the \textit{ArrayLiteral} production in Figure~\ref{fig:array-literal}(a).
Each parser has the \( \code{List[Boolean] => LAParser[T]} \) type
because each production in \( \bnfes \) is parametric with boolean values.
The \( \code{memo} \) is a memoization function for pairs of boolean
parameters and resulting parsers for performance optimization.
The value \( \code{ArrayLiteral} \) corresponds to the
\textit{ArrayLiteral} production.  In the parser, each string literal
such as \( \code{"["} \) or \( \code{"]"} \) denotes a parser for a
terminal symbol.  The \( \code{opt} \) helper function creates
optional parsers.  The parametric non-terminal \textit{ElementList}
with arguments \textsf{Yield} and \textsf{Await} is represented as a
function call \( \code{ElementList(Yield, Await)} \).
The \( \code{\textasciitilde} \) operator combines two parsers
and the \( \code{\^{}\^{}} \) operator describes how to construct ASTs.
When the left-hand side of \( \code{\^{}\^{}} \) is matched, its
right-hand side shows a corresponding AST constructor, where the name
of each constructor has a number denoting the order among alternatives.
For example, the \( \code{ArrayLiteral0} \) constructor corresponds to
the first alternative of the \textit{ArrayLiteral} production.

@@

\textbf{Semantics.}
ECMAScript describes the language semantics as abstract algorithms in English.
While they are written in a natural language, the writing style is
well-organized with ordered steps and tagged tokens.  {\sf Spec Extractor}
reads abstract algorithms with HTML tags and converts them into JSON files.
For example, Figure~\ref{fig:array-literal-eval}(a) presents the
\textbf{Evaluation} abstract algorithm of the third alternative of the
\textit{ArrayLiteral} production in ES10, and it has seven steps.  In the HTML
files describing the abstract algorithms, each non-terminal symbol
(e.g. \textit{ElementList}), local variable (e.g. \textit{array}), code (e.g. \(
\code{"length"} \)), or value (e.g. \textbf{false}) has the \( \code{<nt>} \),
\( \code{<var>} \), \( \code{<code>} \), or \( \code{<emu-val>} \) tag,
respectively.

To translate such abstract algorithms into representations
suitable for manipulation, we define \( \ires \), a specialized
intermediate representation for ECMAScript.  Then,
we develop \textsf{Algorithm Compiler} in Scala using Scala parser
combinators again to convert given abstract algorithms to \( \ires \)
functions.  It also takes \textsf{Compile Rules} as another input,
which has two parts: parsing rules and conversion rules.  They are
\textit{manually} specified for ECMAScript; we explain them in
detail in Section~\ref{sec:compiler}.  Thus, each abstract algorithm is
converted into a function written in \( \ires \) via \textsf{Algorithm Compiler}.
For example, Figure~\ref{fig:array-literal-eval}(b) presents the generated
\( \ires \) function for the \textbf{Evaluation} abstract algorithm shown in
Figure~\ref{fig:array-literal-eval}(a).
The \( \code{ArrayLiteral[2].Evaluation} \) function takes two parameters for
two non-terminal symbols: \( \code{ElementList} \) and \( \code{Elision} \).
The parameter \( \code{Elision} \) has a special value \( \code{absent} \) when
the non-terminal symbol \textit{Elision}$_{\mbox{\footnotesize opt}}$ is not
present.  Thus, we convert the condition in step 4,
``if \textit{Elision} is not present,'' into the equality check with \(
\code{absent} \): \( \code{if (= Elision absent)} \).  Codes are represented as
string values and values are represented as corresponding \( \ires \) values.
For instance, the code \texttt{\small "length"} and the value \textbf{false} are
converted into the string value \( \code{"length"} \) and the boolean value
\( \code{false} \), respectively.

Finally, \( \tool \) constructs {\sf AST-\( \ires \) Translator} with the given
\( \ires \) functions and \textit{manually} specified {\sf Global Setting},
which has minor but necessary information to evaluate JavaScript programs
described in ECMAScript such as the structure of the standard
built-in objects and ECMAScript data types. Putting them all together, we can
translate a given JavaScript program into \( \ires \) via generated {\sf
JavaScript Parser} and {\sf AST-\( \ires \) Translator} by \( \tool \).  Even
though \( \tool \) is not fully automatic because of {\sf Compile Rules} and
{\sf Global Setting}, it could dramatically reduce the efforts to building
parsers and translators from scratch.

In the remainder of this paper, we explain the details of how to
automatically generate parsers (Section~\ref{sec:parser}) and how to compile
abstract algorithms (Section~\ref{sec:compiler}).  After describing implementation
details (Section~\ref{sec:impl}), we evaluate \( \tool \)
(Section~\ref{sec:eval}), discuss related work (Section~\ref{sec:related}), and
conclude (Section~\ref{sec:conclude}).
